---
title: Creating plots
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(flyover)
library(ggplot2)
set.seed(12345)

old <- data.frame(norm  = rnorm(n = 100, mean = 1),
                  exp   = rexp(n = 100, rate = 1),
                  chisq = rchisq(n = 100, df = 5),
                  lnorm = rlnorm(n = 100, meanlog = 0),
                  gamma = rgamma(n = 100, shape = 2),
                  alpha = sample(letters[1:4], size = 100, replace = TRUE, prob = c(1, 2, 3, 4)),
                  hilo  = sample(c("high", "low"), size = 100, replace = TRUE, prob = c(1, 2)),
                  tf    = sample(c(TRUE, FALSE), size = 100, replace = TRUE, prob = c(1, 1)),
                  fruit = sample(c("apple", "banana", "pear"), size = 100, replace = TRUE, prob = c(1, 2, 3)))


new <- data.frame(norm  = rnorm(n = 100, mean = 3),
                  exp   = rexp(n = 100, rate = 3),
                  chisq = rchisq(n = 100, df = 10),
                  lnorm = rlnorm(n = 100, meanlog = 2),
                  gamma = rgamma(n = 100, shape = 4),
                  alpha = sample(letters[1:4], size = 100, replace = TRUE, prob = c(4, 3, 2, 1)),
                  hilo  = sample(c("high", "low"), size = 100, replace = TRUE, prob = c(2, 1)),
                  tf    = sample(c(TRUE, FALSE), size = 100, replace = TRUE, prob = c(1, 3)),
                  fruit = sample(c("apple", "banana", "pear"), size = 100, replace = TRUE, prob = c(3, 2, 1)))


data_list <- enlist_data(old, new, names = c("old data", "new data"))
my_data   <- stack_data(data_list, group_var = "source")
```


## What this article covers

A typical `flyover` workflow has the following steps:

1. Combine different data sets into a single table.
2. Apply a plotting function to the columns of the table.
3. Build a display to navigate the plots.

This vignette will introduce **step 2**: how `flyover`'s plotting functions work and how to easily build comparison plots for different types of data.

This vignette assumes you have a single data set composed of data from different groups for comparison.  We will show how to use that data to build a collection of plots for comparing distributions between groups.


## Data for this example

We simulate a data set with numeric and categorical variables from two different sources (called "old data" and "new data").  Each source has a slightly different data-generating process.  Note the `source` column where these values are stored.  We wish to understand the distributional differences of the other variables between these sources.


```{r view-data}
str(my_data)
```

## Built-in plotting functions

Plotting in `flyover` comes in two flavors, numeric and categorical.  The package comes with basic built-in plotting functions that operate on columns of one of these types, and ignores columns of the other type.  This means you don't have to pre-filter a data frame by the type of data that the plots will use.

* _Numeric_ plotting functions will keep integer and float type variables and ignore everything else.

* _Categorical_ plotting functions will keep character, factor, and logical type variables and will ignore everything else.

Currently, `flyover`'s built-in functions and the data types on which they act are:

```{r data-types, echo = FALSE}
lookup <- flyover:::get_flyover_type_lookup()

lookup_df <- data.frame(function_name = names(lookup),
                        data_type     = lookup,
                        row.names     = NULL)
knitr::kable(lookup_df)
```

Plotting functions ending with `_ridges` are derived from the package `ggridges` ([documentation](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html)) and are well suited to comparing numeric distributions through time, _e.g._ for examining data drift across data pulls occurring on regular intervals.


## Building plots

To build plots, you must specify at least three things:

* a data frame or similar object (such as a `tibble` or `data.table`)
* the `flyover` plotting function you wish to use
* the name of the variable that distinguishes groups (defaults to `flyover_id_` which is the default output of `stack_data`).

For each variable of the type associated with your chosen plotting function, `flyover` will generate a plot.  However, these plots are not printed yet; they are merely stored in a tibble until they are used to build displays.


### Numeric plots

As an example, suppose we wish to compare the numeric variables in our example data across the values of the group variable `source`.  We are interested in histograms.  Then we would build the plots in the following manner:

```{r make-hist}
my_histograms <- build_plots(my_data, flyover_histogram, group_var = "source")
my_histograms
```

Notice that the output is a tibble having column `plot` which contains one plot for each numeric variable in the data.  (See below for a discussion on the `cogs` columns.)  The function does not print the plots directly at this stage.  However, the individual plots can still be accessed:

```{r print-hist, message = FALSE}
my_histograms$plot[[1]]
```


You are not limited to the default appearance of the built-in plotting functions.  You can modify these plots by passing additional `ggplot2` elements as a list:

```{r modify-hist, message = FALSE}
my_histograms_mod <- build_plots(my_data, flyover_histogram, group_var = "source",
                                 plot_mods = list(xlab("new X-axis label"),
                                                  theme_classic(),
                                                  scale_fill_viridis_d()))

my_histograms_mod$plot[[1]]
```


### Categorical plots

In a similar way, we can compare distributions of categorical variables across groups by examining their proportions using bar plots.

```{r make-bar}
my_bars <- build_plots(my_data, flyover_bar_fill, group_var = "source")
my_bars
```

As before, you can view individual plots by calling them from this tibble.

```{r print-bar, message = FALSE}
my_bars$plot[[1]]
```


## Cognostics

The tibble returned by `build_plots` contains a column called `cogs` which is short for "cognostics".  You can read more about cognostics in the `trelliscopejs` [documentation](https://hafen.github.io/trelliscopejs/articles/trelliscopejs.html). In short, they are pieces of metadata that are associated with each plot in order to sort and filter plots.  They are stored as tibbles, and for numeric plots they look like this:

```{r}
str(my_histograms$cogs[[1]])
```

These are summaries of features of the data used to generate the first plot.  They represent the largest differences (in percentage terms) between any groups for that variable.

For categorical plots, the cognostics look like this:

```{r}
str(my_bars$cogs[[1]])
```

The next article will clarify how cognostics are used in trelliscope displays.


## Using a pipe

Notice that these functions -- indeed, all the functions in this package -- are data-first, meaning they are pipe-friendly.  Thus you could in theory write a pipeline like this:

```
my_histograms <-
  old %>%
  enlist_data(new, names = c("old data", "new data")) %>%
  stack_data() %>%
  build_plots(flyover_histogram)
```