---
title: Using custom plots
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using custom plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(flyover)
set.seed(12345)

old <- data.frame(norm  = rnorm(n = 100, mean = 1),
                  exp   = rexp(n = 100, rate = 1),
                  chisq = rchisq(n = 100, df = 5),
                  lnorm = rlnorm(n = 100, meanlog = 0),
                  gamma = rgamma(n = 100, shape = 2),
                  alpha = sample(letters[1:4], size = 100, replace = TRUE, prob = c(1, 2, 3, 4)),
                  hilo  = sample(c("high", "low"), size = 100, replace = TRUE, prob = c(1, 2)),
                  tf    = sample(c(TRUE, FALSE), size = 100, replace = TRUE, prob = c(1, 1)),
                  fruit = sample(c("apple", "banana", "pear"), size = 100, replace = TRUE, prob = c(1, 2, 3)))


new <- data.frame(norm  = rnorm(n = 100, mean = 3),
                  exp   = rexp(n = 100, rate = 3),
                  chisq = rchisq(n = 100, df = 10),
                  lnorm = rlnorm(n = 100, meanlog = 2),
                  gamma = rgamma(n = 100, shape = 4),
                  alpha = sample(letters[1:4], size = 100, replace = TRUE, prob = c(4, 3, 2, 1)),
                  hilo  = sample(c("high", "low"), size = 100, replace = TRUE, prob = c(2, 1)),
                  tf    = sample(c(TRUE, FALSE), size = 100, replace = TRUE, prob = c(1, 3)),
                  fruit = sample(c("apple", "banana", "pear"), size = 100, replace = TRUE, prob = c(3, 2, 1)))


data_list <- enlist_data(old, new)
my_data   <- stack_data(data_list, group_var = "source")
```


## What this article covers

A typical `flyover` workflow has the following steps:

1. Combine different data sets into a single table.
2. Apply a plotting function to the columns of the table.
3. Build a display to navigate the plots.

This vignette will introduce **alternative plotting functions** for step 2.  While there are some built-in `flyover` plotting functions ready to use, there will be times that you need something customized.  You can write your own plotting function to pass into the `build_plots` function.


## Data for this example

The data used in this example is the same as in the previous vignettes.


```{r view-data}
str(my_data)
```


## Structure of `flyover` plotting functions

The built-in plotting functions that ship with this package are very simple wrappers for `ggplot2` plots.  For example, this is the body of the function `flyover_histogram`:

```{r fun-body, echo = FALSE}
flyover_histogram
```


Important features to note are the following:

* the first argument is the `tbl` containing all the data to plot;
* the second and third arguments are character strings specifying the variable to plot and the grouping variable, respectively;
* the dots (`...`) provide further modification to the plot.  These modifications can be specified in the call to `build_plots` and are passed down to these plotting functions;
* the plotting is performed by `ggplot2`.

When you write a custom plotting function, this is the general structure you should follow for compatibility with the workings of `build_plots`.  Of course, you are free to use any themes or other customizations you like.


## Defining a custom plot

Let's say we wish to compare distributions not through the use of histograms, but with empirical cumulative distribution functions (ECDFs).  After some research on the `ggplot2` [website](https://ggplot2.tidyverse.org/index.html), we find the right syntax.

We can define the function for use in `flyover` as follows:

```{r ecdf-definition}
custom_ecdf <- function(tbl, var, group_var, ...) {
  ggplot(tbl, aes_string(x = var, color = group_var)) +
    stat_ecdf(size = 2, ...) +
    theme_minimal(base_size = 14)
}
```


## Using a custom plot

We can now pass the custom plot function to `build_plots` in the same way as a built-in `flyover` function, _with one important addition_.  The `build_plots` function doesn't have any knowledge about what kind of data our custom plot operates on.  We need to supply this knowledge via the `keep_type` argument, which tells `build_plots` which type of variables to keep before plotting.  You must pass either `"numeric"` or `"categorical"`.  Since `custom_ecdf` operates on numeric data, that's what we choose.

```{r ecdf-display, eval = FALSE}
my_ecdf <- build_plots(my_data, custom_ecdf, "source", keep_type = "numeric")
build_display(my_ecdf,
              display_name = "custom ECDF",
              output_dir   = "display-ecdf")
```

<iframe src="display-ecdf/" height=500 width=850 scrolling="no" frameborder="0" webkitallowfullscreen="true" mozallowfullscreen="true" allowfullscreen="true"></iframe>